Impact of Roe v. Wade’s Overturn: Abortion Restrictions and Infant Mortality in the U.S. Using a Difference-in-Differences Analysis

### 2024-11-29 paper/ExaminingtheimpactOverturnofRoev.Wade:BanningofAbortiononInfantMortalityRatesintheUnitedStatesUsingaDifference-in-DifferencesApproach.qmd
update paper with format error

-   **Treatment Group Indicator `abortion_illegal`**: A binary variable
    indicating the group: 1 if abortion is illegal in the state. 0 if
    abortion remains legal.

-   **Time Indicator `after_injunction`**: A binary variable indicating
    the time period: 0 for the period before the injunction was
    implemented. 1 for the period after the injunction was implemented.

-   **Interaction Term`after_injunction` × `abortion_illegal`**: The
    interaction term captures the additional change in the outcome
    (e.g., percentage of support, behavior, or health-related metrics)
    for regions where abortion is illegal after the injunction
    implemented, beyond any changes observed in regions where abortion
    remains legal.

@@

-   **`age_of_mother`**: The age of the mother at the time of birth,
    represented as a continuous variable.

-   **`mothers_single_race`**: Self-identified race of the mother,
    represented as a categorical variable.

-   **`year_of_death`**: The year in which the infant death occurred,
    represented as a continuous variable.

@@

-   **`state`**: Mother's residence state when the infant was born as a
    random effect in the model.

@@

One model includes control variables— `age_of_mother`,
`mothers_single_race`, and `year_of_death` to adjust for confounders
that may influence infant death rates. This adjustment addresses
compositional differences assumed in the DID framework and ensures
unbiased estimates of the intervention’s effect. The second model
excludes these controls to focus on the overall intervention impact,
avoid overfitting, and prioritize simplicity where variability is
adequately accounted for by design or random effects. By fitting two
models, we compare outcomes to refine estimates and evaluate the direct
impact of the intervention.

>>>>

-   **Treatment Group Indicator (`abortion_illegal`)**: A binary variable
    indicating the group: 1 if abortion is illegal in the state. 0 if
    abortion remains legal.

-   **Time Indicator (`after_injunction`)**: A binary variable indicating
    the time period: 0 for the period before the injunction was
    implemented. 1 for the period after the injunction was implemented.

-   **Interaction Term (`after_injunction` × `abortion_illegal`)**: The
    interaction term captures the additional change in the outcome
    (e.g., percentage of support, behavior, or health-related metrics)
    for regions where abortion is illegal after the injunction
    implemented, beyond any changes observed in regions where abortion
    remains legal.

@@

-   **Age of Mother (`age_of_mother`)**: The age of the mother at the time of birth,
    represented as a continuous variable.

-   **Mother's Single Race (`mothers_single_race`)**: Self-identified race of the mother,
    represented as a categorical variable.

-   **Year of Death (`year_of_death`)**: The year in which the infant death occurred,
    represented as a continuous variable.

@@

-   **State (`state`)**: Mother's residence state when the infant was born as a
    random effect in the model.

@@

One model includes control variables — Age of Mother,
Mother's Single Race, and Year of Death to adjust for confounders
that may influence infant death rates. This adjustment addresses
compositional differences assumed in the DID framework and ensures
unbiased estimates of the intervention’s effect. The second model
excludes these controls to focus on the overall intervention impact,
avoid overfitting, and prioritize simplicity where variability is
adequately accounted for by design or random effects. By fitting two
models, we compare outcomes to refine estimates and evaluate the direct
impact of the intervention.
